---
title: "Quality of Activity in Fitness"
subtitle: "Final Project Report - Practical Machine Learning Course"
author: Carlos A. Giménez
output:
  html_document: default
---

# Introduction

The goal of this project is to predict the manner in which a group of people did exercises, these 
are based on a previus captured information in sensors of movement.

The data set are divided in to two parts Training and Test. From the Training data set, we take 
the "classe" variable like target of prediction.

## Background

The follow background of the project was extracted from the Practical Machine Learning Course
of the Johns Hoopsking University:

"Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large 
amount of data about personal activity relatively inexpensively. These type of devices are part of 
the quantified self movement – a group of enthusiasts who take measurements about themselves 
regularly to improve their health, to find patterns in their behavior, or because they are tech 
geeks. One thing that people regularly do is quantify how much of a particular activity they do, 
but they rarely quantify how well they do it. In this project, your goal will be to use data from
accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform
barbell lifts correctly and incorrectly in 5 different ways. More information is available from the
website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting 
Exercise Dataset)."

## Session information
For this project the sistem was:
```{r}
sessionInfo()
```

# Getting and Cleaning Data

The data sources was:

For training data

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

For test data

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

The follow step consist in load the data:
```{r}
if(!file.exists("./data")){dir.create("./data")}

training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```
Let`s take a look to datasets
```{r}
dim(training); dim(testing)
```
Let`s check the target variable
```{r}
summary(training$classe)
```
As we can see, the "classe" variable contain five differents category, 

## Preparing dataset

```{r}
set.seed(12345)
library(caret)
```

```{r}
inTrain <- createDataPartition(y=training$classe, p=0.7, list=FALSE)
Training_wk <- training[inTrain, ]
Testing_wk <- training[-inTrain, ]
dim(Training_wk); dim(Testing_wk)
```

## Cleaning Stage

The first step is handle values near to zero

```{r}
nzv <- nearZeroVar(Training_wk, saveMetrics=TRUE)
Training_wk <- Training_wk[,nzv$nzv==FALSE]

nzv <- nearZeroVar(Testing_wk,saveMetrics=TRUE)
Testing_wk <- Testing_wk[,nzv$nzv==FALSE]
```


The follow steps consist in cleaning the data. First, we need check the "NA" values, if we have
many NAs in one column we need discard that coulumn

```{r}
clean_train <- Training_wk
for(i in 1:length(Training_wk)) { if( sum(is.na( Training_wk[, i] ) ) /nrow(Training_wk) >= .5) {
        for(j in 1:length(clean_train)) {
            if( length( grep(names(Training_wk[i]), names(clean_train)[j]) ) == 1)  {
                clean_train <- clean_train[ , -j]}}}}
Training_wk <- clean_train
```

```{r}
dim(Training_wk)
```

```{r}
clean1 <- colnames(Training_wk)
clean2 <- colnames(Training_wk[, -59]) 
```
```{r}
# Extract only columns that we have in Training_wk
Testing_wk <- Testing_wk[clean1]         
testing <- testing[clean2]
```

```{r}
# To get the same class 
testing <- rbind(Training_wk[2, -59] , testing)
testing <- testing[-1,]
```


# Prediction

## Decision Tree
```{r}
require(rpart)
```

```{r}
DT <- rpart(classe ~ ., data=Training_wk, method="class")
```
```{r}
prediction <- predict(DT, Testing_wk, type = "class")
matrix_cm <- confusionMatrix(prediction, Testing_wk$classe)

matrix_cm$overall
```


## Random Forest

```{r}
require(randomForest)
```


```{r}
RF <- randomForest(classe ~ ., data = Training_wk)
prediction_RF <- predict(RF, Testing_wk, type = "class")
confmat <- confusionMatrix(prediction_RF, Testing_wk$classe)
```

## Apply the best model

The best model is Random Forest,

```{r}
dim(testing); dim(Testing_wk)
```



```{r}
final <- predict(RF, testing, type = "class")
final
```

# Conclusion
The randomForest approach had much higher predictive accuracy than the decision tree classifier. For
this project in particular.